The Effect of Different Writing Tasks on Linguistic Style:\\ A Case Study of the ROC Story Cloze Task

### 2017-04-18 docs/language_constraint_conll.tex
CoNLL V2 (Roy)

In order to further validate our method, we also directly tackle the
story cloze task\nascomment{cite the challenge?  a reader may
  not realize that there is a shared task on this}\resolved{\roy{Maybe ``challenge'' was not the right term here. We basically mean the story cloze task. The shared task was the original task w/o any modification, so I am not referring to it specifically here}}. 
 Adapting our classifier to the task, we obtain 72.4\% accuracy, \com{a
 12.5\% increase over the previously reported state-of-the-art  \cite{Salle:2016}}\roy{only 2\% below state-of-the-art results}.
We also show that the style differences captured by our model can be combined with neural language models to make a better use of the story context. 
Our final model that combines context with stylistic features achieves \roy{new state-of-the-art results---}
75.2\%---an additional 2.8\% gain\com{, 15.3\% better than the best
published result}.

@@

The ROC story corpus consists of 49,255 five-sentence 
stories, collected on Amazon Mechanical Turk (AMT).\footnote{Recently,
  an additional 53K stories were released, which results in roughly
  100K stories.}
Workers were instructed to write a coherent self-contained story, which has a clear beginning and end. 
To collect a broad spectrum of commonsense knowledge, there was no imposed subject for the stories,
which resulted in a wide range of different topics.

@@

\isection{Surface Analysis of the Story Cloze Task}{Surface}
We begin by computing several characteristics of the three types of endings: {\it original} endings (from the ROC story corpus training set), {\it right} endings and {\it wrong} endings (both from the story cloze task development set).
Our analysis  reveals several style differences between different groups. 
First, {\it original} endings are on average longer (11 words per
sentence) than {\it right} endings (8.75 words), which are in turn
slightly longer than {\it wrong} ones (8.47 words). 
Previous work has shown that sentence length is also indicative of whether a text was deceptive \cite{yancheva2013automatic,qin2004exploratory}. 
Although writing {\it wrong} sentences is not the same as lying, it is not entirely surprising to observe similar trends in both tasks.
%Furthermore, \citet{Buller:1996} suggests lying increases one's cognitive load, which in turn could affect sentence length.
%\yc{Is there psycholinguistic literature that we can cite that may have shown that cognitive burden causes writers to write shorter sentences...?}

@@

\begin{figure*}[t!]
\begin{small}
\subfloat[POS tags]{
%\begin{subfigure}{.5\textwidth}\center
\begin{tikzpicture}
    \begin{axis}[
        width  = 0.4*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=\pgflinewidth,
        x=1.25cm,
       bar width=5pt,
        ymajorgrids = true,
        ylabel = {Frequency in Corpus (\%)},
%        symbolic x coords={{\it original}, {\it right}, {\it wrong}},
        symbolic x coords={NN, VBD, PRP, DT, NNP},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=3,
        legend cell align=left,
        legend style={
        		legend columns=-1,
                at={(1,1.05)},
                anchor=south east,
                column sep=1ex
        }
    ]
    \addplot[style={blue,fill=blue,mark=none}]
            coordinates {(NN, 15.3) (VBD, 13.4) (PRP, 10.1) (DT, 9.1) (NNP, 4.7)};

@@

%        \legend{NN, VBD, PRP, DT, NNP}
	\legend{{\it original}, {\it right}, {\it wrong}}
    \end{axis}
\end{tikzpicture}
%\caption{
\label{roc_pos_distribution}
%}
%\end{subfigure}
}
\qquad\qquad\qquad
\subfloat[Words]{%\begin{subfigure}{.5\textwidth}\center
\begin{tikzpicture}
    \begin{axis}[
        width  = 0.4*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=\pgflinewidth,
        x=1.25cm,
       bar width=5pt,
        ymajorgrids = true,
%        ylabel = {Frequency in Corpus (\%)},
    %    symbolic x coords={{\it original}, {\it right}, {\it wrong}},
    symbolic x coords={to, and, I, hates, !},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=0,
        legend cell align=left,
        legend style={
        		legend columns=-1,
                at={(1,1.05)},
                anchor=south east,
                column sep=1ex
        }
    ]

@@

        \addplot[style={gray,fill=gray,mark=none}]
            coordinates {({\it original}, 0.9) ({\it right}, 0.2) ({\it wrong}, 0.1)};
}
%        \legend{to, and, I, hates, !}
        	\legend{{\it original}, {\it right}, {\it wrong}}

    \end{axis}
\end{tikzpicture}
%\caption{
\label{roc_word_distribution}
%Words}
%\end{subfigure}
}
\end{small}
\caption{The distribution of five frequent POS tags
  (\ref{roc_pos_distribution}) and words (\ref{roc_word_distribution})
  across \emph{original} endings (story cloze training set), and
  \emph{right} and \emph{wrong} endings (from the story cloze task).}

@@

We train a logistic regression classifier to categorize an ending,
either as {\it right} vs.~{\it wrong} or as {\it original} vs.~new ({\it right}).
Each feature vector is computed using the words in one ending, without considering earlier parts of the story. 
We use the following style features.

@@

\isectionb{Experiments}
We design two experiments to answer our research questions. 
The first is an attempt to distinguish between {\it right} and {\it wrong} endings,
the second  between {\it original} endings and new ({\it right}) endings.
For completeness, we also run a third experiment, which compares between {\it original} and {\it wrong} endings. 
We describe all experiments below.

@@

Here the goal is to measure whether writing the ending as part of a
story imposes different style compared to writing a {\it new} ({\it right}) ending to an existing story.
We use the endings of the ROC stories as our {\it original} samples and {\it right} endings from the story cloze task   \resolved{\nascomment{if you're using the test examples as
  training data, what are you using for test? and development?  I would have thought
  you'd follow the same pattern as above ... }}as {\it new} samples.
As there are far more {\it original} instances than {\it new}
instances, we randomly select five  {\it original} sets, each with the same number of
instances  \resolved{\nascomment{(give the number in parentheses, for train/dev/test)} }as we have
\emph{new} instances (3,366 training endings, 374 development endings, and 3,742 test endings).
%We randomly sample 5 {\it original} sets and repeat the classification experiments.
We train five classifiers with each of the {\it original} training sets, and report the average classification result.

@@

 \resolved{
\begin{figure}
\begin{tikzpicture}
\small
    \begin{axis}[
        width  = 0.45*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=2*\pgflinewidth,
       % bar width=14pt,
        ymajorgrids = true,
        ylabel = {Classification Performance},
        symbolic x coords={{\it Correct/Wrong}, {\it Original/New}, {\it Original/Wrong}},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=0.4,
        legend cell align=left,
        legend style={
                at={(1,1.05)},
                anchor=south east,
                column sep=1ex
        }
    ]
        \addplot[style={bblue,fill=bblue,mark=none}]
            coordinates {({\it Correct/Wrong}, 0.5) ({\it Original/New}, 0.5) ({\it Original/Wrong},.5)};

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|l|r|} \hline
{\bf Experiment} & {\bf Accuracy} \\ \hline
{\sl right vs. wrong} & 0.645 \\ \hline
{\sl original vs. right} & 0.685 \\ \hline
{\sl original vs. wrong} & 0.756 \\ \hline
\end{tabular}
\end{center}
\caption{\label{results}Results of  experiments 1 ({\it right vs.~wrong}), 2 ({\it original vs.~right (new)}) and 3 ({\it
  original vs.~wrong (new)} endings).
In all cases, our setup implies a 50\% random baseline.}
\end{table}

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|c|c|} \hline
{\bf Feature Type} & {\bf Accuracy}\\ \hline
Word $n$-grams & 0.612 \\ \hline
Character $n$-grams & 0.639 \\ \hline
Full model & 0.645 \\ \hline

@@

Second, {\it right} endings will make wider use of coordination structures, as well as adjectives.
The latter might indicate that writing coherent stories inspires the authors to write more descriptive text compared to incoherent ones, 
as is the case in truthful vs.~deceptive text \cite{ott2011finding}.
Finally, we notice a few syntactic differences: {\it right} endings will more often use infinite verb structure, while {\it wrong} endings prefer gerunds (VBG).

@@

\end{tabular}
\label{exp2_features}
}
\caption{
The top 5 most heavily weighted features for predicting {\it right} vs.~{\it wrong} endings (\ref{exp1_features}) and  {\it original} vs.~new ({\it right}) endings (\ref{exp2_features}). 
{\it length} is the sentence length feature (see \secref{Model}).
}
%\end{table}

@@

Finally, when comparing the two groups of salient features from both experiments, we find an interesting trend.
Several features, such as ``\textsc{start} NNP'' and  ``NN .'', which indicate {\it wrong} sentences in Experiment 1, are used to predict new (i.e., {\it right}) endings in Experiment 2. 
This indicates that, for instance, incoherent endings have a stronger tendency to begin with a proper noun compared to coherent endings, 
which in turn are more likely to do so than original endings. 
This partially explains why distinguishing between {\it original} and {\it wrong} endings is an easier task compared to the other pairs (\secref{Results}).

@@

Applying our story cloze classifier to this dataset yields 53.2\% classification accuracy---higher than a random baseline, but substantially lower than our story cloze task results. 
While this could be partially explained by the smaller data size of the COPA dataset (1,000 samples compared to 3,742 in the story cloze task), this indicates that simple instructions may help alleviate the effects of writing style found in this paper.

Another way to avoid such effects is to use naturally occurring text, as used in recent machine reading tasks (see \secref{Related}). 
%However,  these approaches have their shortcomings, and do not capture the same signal as intended by the author of the story cloze task.
Finally, a different approach is to have people rate sentences from naturally occurring text by parameters
such as coherence (or, very differently, the level of surprise),
rather than asking them to generate new text.

>>>>

In order to further validate our method, we also directly tackle the
story cloze task. 
Adapting our classifier to the task, we obtain 72.4\% accuracy, \com{a
 12.5\% increase over the previously reported state-of-the-art  \cite{Salle:2016}}\roy{only 2\% below state-of-the-art results}.
We also show that the style differences captured by our model can be combined with neural language models to make a better use of the story context. 
Our final model that combines context with stylistic features achieves \roy{new state-of-the-art results---}
75.2\%---an additional 2.8\% gain\com{, 15.3\% better than the best
published result}.

@@

The ROC story corpus consists of 49,255 five-sentence 
stories, collected on Amazon Mechanical Turk (AMT).\footnote{Recently,
  additional 53K stories were released, which results in roughly
  100K stories.}
Workers were instructed to write a coherent self-contained story, which has a clear beginning and end. 
To collect a broad spectrum of commonsense knowledge, there was no imposed subject for the stories,
which resulted in a wide range of different topics.

@@

\isection{Surface Analysis of the Story Cloze Task}{Surface}
We begin by computing several characteristics of the three types of endings: {\it original} endings (from the ROC story corpus training set), {\it right} endings and {\it wrong} endings (both from the story cloze task development set).
Our analysis  reveals several style differences between different groups. 
First, {\it original} endings are on average longer (11 words per
sentence) than {\it right} endings (8.75 words), which are in turn
slightly longer than {\it wrong} ones (8.47 words). 
The latter finding is consistent with previous work, which has shown that sentence length is also indicative of whether a text was deceptive \cite{yancheva2013automatic,qin2004exploratory}. 
Although writing {\it wrong} sentences is not the same as lying, it is not entirely surprising to observe similar trends in both tasks.
%Furthermore, \citet{Buller:1996} suggests lying increases one's cognitive load, which in turn could affect sentence length.
%\yc{Is there psycholinguistic literature that we can cite that may have shown that cognitive burden causes writers to write shorter sentences...?}

@@

\begin{figure*}[t!]
\begin{small}
\subfloat[POS tags]{
%\begin{subfigure}{.5\textwidth}\center
\begin{tikzpicture}
    \begin{axis}[
        width  = 0.4*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=\pgflinewidth,
        x=1.25cm,
       bar width=5pt,
        ymajorgrids = true,
        ylabel = {Frequency in Corpus (\%)},
%        symbolic x coords={{\it original}, {\it right}, {\it wrong}},
        symbolic x coords={NN, VBD, PRP, DT, NNP},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=3,
%        legend cell align=left,
%        legend style={
%        		legend columns=-1,
%                at={(1,1.05)},
%                anchor=south east,
%                column sep=1ex
%        }
    ]
    \addplot[style={blue,fill=blue,mark=none}]
            coordinates {(NN, 15.3) (VBD, 13.4) (PRP, 10.1) (DT, 9.1) (NNP, 4.7)};

@@

%        \legend{NN, VBD, PRP, DT, NNP}
%	\legend{{\it original}, {\it right}, {\it wrong}}
    \end{axis}
\end{tikzpicture}
%\caption{
\label{roc_pos_distribution}
%}
%\end{subfigure}
}
\qquad\qquad\qquad
\subfloat[Words]{%\begin{subfigure}{.5\textwidth}\center
\begin{tikzpicture}
    \begin{axis}[
        width  = 0.4*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=\pgflinewidth,
        x=1.25cm,
       bar width=5pt,
        ymajorgrids = true,
%        ylabel = {Frequency in Corpus (\%)},
    %    symbolic x coords={{\it original}, {\it right}, {\it wrong}},
    symbolic x coords={to, and, I, hates, !},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=0,
%        legend cell align=left,
%        legend style={
%        		legend columns=-1,
%                at={(1,1.05)},
%                anchor=south east,
%                column sep=1ex
%        }
    ]

@@

        \addplot[style={gray,fill=gray,mark=none}]
            coordinates {({\it original}, 0.9) ({\it right}, 0.2) ({\it wrong}, 0.1)};
}
%        \legend{to, and, I, hates, !}
%        	\legend{{\it original}, {\it right}, {\it wrong}}

    \end{axis}
\end{tikzpicture}
%\caption{
\label{roc_word_distribution}
%Words}
%\end{subfigure}
}
\end{small}
\caption{The distribution of five frequent POS tags
  (\ref{roc_pos_distribution}) and words (\ref{roc_word_distribution})
  across {\color{blue}{\emph{original}}} endings (story cloze training set), and
  {\color{forestgreen}{\emph{right}}} and {\color{red}{\emph{wrong}}} endings (from the story cloze task).}

@@

We train a logistic regression classifier to categorize an ending,
either as {\it right} vs.~{\it wrong} or as {\it original} vs.~{\it new} ({\it right}).
Each feature vector is computed using the words in one ending, without considering earlier parts of the story. 
We use the following style features.

@@

\isectionb{Experiments}
We design two experiments to answer our research questions. 
The first is an attempt to distinguish between {\it right} and {\it wrong} endings,
the second  between {\it original} endings and {\it new} ({\it right}) endings.
For completeness, we also run a third experiment, which compares between {\it original} and {\it wrong} endings. 
We describe all experiments below.

@@

Here the goal is to measure whether writing the ending as part of a
story imposes different style compared to writing a {\it new} ({\it right}) ending to an existing story.
We use the endings of the ROC stories as our {\it original} samples and {\it right} endings from the story cloze task   \resolved{\nascomment{if you're using the test examples as
  training data, what are you using for test? and development?  I would have thought
  you'd follow the same pattern as above ... }}as {\it new} samples.
As there are far more {\it original} instances than {\it new}
instances, we randomly select five  {\it original} sets, each with the same number of
instances  \resolved{\nascomment{(give the number in parentheses, for train/dev/test)} }as we have
\emph{new} instances (3,366 training endings, 374 development endings, and 3,742 test endings).
%We randomly sample 5 {\it original} sets and repeat the classification experiments.
We train five classifiers, one with each of the {\it original} training sets, and report the average classification result.

@@

 \resolved{
\begin{figure}
\begin{tikzpicture}
\small
    \begin{axis}[
        width  = 0.45*\textwidth,
        height = 4cm,
      %  major x tick style = transparent,
        ybar=2*\pgflinewidth,
       % bar width=14pt,
        ymajorgrids = true,
        ylabel = {Classification Performance},
        symbolic x coords={{\it Correct/Wrong}, {\it Original/New}, {\it Original/Wrong}},
        xtick = data,
        scaled y ticks = false,
        %enlarge x limits=0.25,
        ymin=0.4,
%        legend cell align=left,
%        legend style={
%                at={(1,1.05)},
%                anchor=south east,
%                column sep=1ex
%        }
    ]
        \addplot[style={bblue,fill=bblue,mark=none}]
            coordinates {({\it Correct/Wrong}, 0.5) ({\it Original/New}, 0.5) ({\it Original/Wrong},.5)};

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|l|c|} \hline
{\bf Experiment} & {\bf Accuracy} \\ \hline
{\sl right vs. wrong} & 0.645 \\ \hline
{\sl original vs. right} & 0.685 \\ \hline
{\sl original vs. wrong} & 0.756 \\ \hline
\end{tabular}
\end{center}
\caption{\label{results}Results of  experiments 1 ({\it right vs.~wrong}), 2 ({\it original vs.~right (new)}) and 3 ({\it
  original vs.~wrong (new)} endings).
In all cases, our setup implies a 50\% random baseline.}
\end{table}

@@

\begin{table}[!t]
\begin{center}
\begin{tabular}{|l|c|} \hline
{\bf Feature Type} & {\bf Accuracy}\\ \hline
Word $n$-grams & 0.612 \\ \hline
Character $n$-grams & 0.639 \\ \hline
Full model & 0.645 \\ \hline

@@

Second, {\it right} endings make wider use of coordination structures, as well as adjectives.
The latter might indicate that writing coherent stories inspires the authors to write more descriptive text compared to incoherent ones, 
as is the case in truthful vs.~deceptive text \cite{ott2011finding}.
Finally, we notice a few syntactic differences: {\it right} endings more often use infinite verb structure, while {\it wrong} endings prefer gerunds (VBG).

@@

\end{tabular}
\label{exp2_features}
}
\caption{
The top 5 most heavily weighted features for predicting {\it right} vs.~{\it wrong} endings (\ref{exp1_features}) and  {\it original} vs.~{\it new} ({\it right}) endings (\ref{exp2_features}). 
{\it length} is the sentence length feature (see \secref{Model}).
}
%\end{table}

@@

Finally, when comparing the two groups of salient features from both experiments, we find an interesting trend.
Several features, such as ``\textsc{start} NNP'' and  ``NN .'', which indicate {\it wrong} sentences in Experiment 1, are used to predict {\it new} (i.e., {\it right}) endings in Experiment 2. 
This indicates that, for instance, incoherent endings have a stronger tendency to begin with a proper noun compared to coherent endings, 
which in turn are more likely to do so than original endings. 
This partially explains why distinguishing between {\it original} and {\it wrong} endings is an easier task compared to the other pairs (\secref{Results}).

@@

Applying our story cloze classifier to this dataset yields 53.2\% classification accuracy---higher than a random baseline, but substantially lower than our story cloze task results. 
While this could be partially explained by the smaller data size of the COPA dataset (1,000 samples compared to 3,742 in the story cloze task), this indicates that simple instructions may help alleviate the effects of writing style found in this paper.
Another way to avoid such effects is to have people rate naturally occurring sentences  by parameters
such as coherence (or, conversely, the level of surprise),
rather than asking them to generate new text.
